---
title: "Forecasting Global Search Interest in Artificial Intelligence"
subtitle: "MATH1318 Time Series Analysis — Final Assignment"
author: |
  | Jayasri Ravichandran 
date: "June 2026"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    number_sections: true
    df_print: kable
  pdf_document:
    toc: true
    number_sections: true
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,
  message   = FALSE,
  warning   = FALSE,
  fig.width = 10,
  fig.height = 6,
  fig.align = "center"
)
```

---

**Research Question:** What are the most accurate 10-month forecasts for global AI search interest (Google Trends) for the period June 2026 to March 2027?

**Data:** Google Trends monthly search interest index for "artificial intelligence" worldwide, January 2004 – May 2026 (n = 269).

---


## Libraries and Helper Functions

```{r setup, message=FALSE, warning=FALSE, echo=TRUE}
# ---- 0. LIBRARIES & HELPER FUNCTIONS ----------------------------------------

rm(list = ls())

library(TSA)
library(fUnitRoots)
library(forecast)
library(lmtest)
library(tseries)
library(fGarch)
library(rugarch)
library(Hmisc)

# --- sort.score() ---
sort.score <- function(x, score = c("bic", "aic")) {
  if (score == "aic") {
    x[with(x, order(AIC)), ]
  } else if (score == "bic") {
    x[with(x, order(BIC)), ]
  } else {
    warning('score = "x" only accepts valid arguments ("aic","bic")')
  }
}

# --- safe.AIC() / safe.BIC() — handles CSS-fallback models -------------------
get_ic <- function(model, type = "AIC") {
  val <- tryCatch(
    if (type == "AIC") AIC(model) else BIC(model),
    error = function(e) numeric(0)
  )
  if (length(val) == 0 || is.na(val)) return(NA_real_)
  as.numeric(val)
}

safe.AIC <- function(...) {
  models <- list(...)
  nms    <- sapply(match.call()[-1], deparse)
  vals   <- vapply(models, get_ic, numeric(1), type = "AIC")
  result <- data.frame(AIC = vals, row.names = nms)
  result[order(result$AIC, na.last = TRUE), , drop = FALSE]
}
safe.BIC <- function(...) {
  models <- list(...)
  nms    <- sapply(match.call()[-1], deparse)
  vals   <- vapply(models, get_ic, numeric(1), type = "BIC")
  result <- data.frame(BIC = vals, row.names = nms)
  result[order(result$BIC, na.last = TRUE), , drop = FALSE]
}

# --- ljung.box.plot() — replaces FitAR::LBQPlot, no extra package needed -----
ljung.box.plot <- function(res, lag.max = 30) {
  lags  <- 1:lag.max
  pvals <- sapply(lags, function(i) Box.test(res, lag = i, type = "Ljung-Box")$p.value)
  plot(lags, pvals, xlab = "Lag", ylab = "p-value", ylim = c(0, 1),
       main = "Ljung-Box Q-test p-values",
       pch = 20, col = ifelse(pvals < 0.05, "red", "steelblue"))
  abline(h = 0.05, col = "red", lty = 2)
  legend("topright", legend = c("p-value", "5% level"),
         col = c("steelblue", "red"), pch = c(20, NA), lty = c(NA, 2))
}

# --- residual.analysis() -----------------------------------------------------
residual.analysis <- function(model, std = TRUE, start = 2, lag.max = 30,
                               class = c("ARIMA", "SARIMA", "GARCH",
                                         "ARMA-GARCH", "garch", "fGARCH")[1]) {
  library(TSA)
  if ((class == "ARIMA") | (class == "SARIMA")) {
    res.model <- if (std) rstandard(model) else residuals(model)
  } else if (class == "GARCH") {
    res.model <- model$residuals[start:model$n.used]
  } else if (class == "garch") {
    res.model <- model$residuals[start:model$n.used]
  } else if (class == "ARMA-GARCH") {
    res.model <- model@fit$residuals
  } else if (class == "fGARCH") {
    res.model <- model@residuals
  } else {
    stop("'class' must be one of: ARIMA, SARIMA, GARCH, garch, ARMA-GARCH, fGARCH")
  }
  res.plain <- as.numeric(res.model)   # fix fractional lag bug in ACF
  layout(matrix(c(1,2,3,4,5,6), nrow = 3, ncol = 2, byrow = TRUE))
  par(mar = c(4, 4, 3, 1))
  plot(res.plain, type = "o", ylab = "Standardised Residuals",
       main = "Time Series Plot of Standardised Residuals")
  abline(h = 0)
  hist(res.plain, main = "Histogram of Standardised Residuals",
       xlab = "Standardised Residuals")
  qqnorm(res.plain, main = "QQ Plot of Standardised Residuals")
  qqline(res.plain, col = 2)
  if (class == "SARIMA") {
    seasonal_acf(res.plain, main = "ACF of Standardised Residuals")
  } else {
    acf(res.plain, main = "ACF of Standardised Residuals")
  }
  print(shapiro.test(res.plain))
  ljung.box.plot(res.plain, lag.max = lag.max)
  layout(1)
  par(mar = c(5, 4, 4, 2) + 0.1)
}

# --- seasonal_acf() / seasonal_pacf() ---
helper <- function(class = c("acf", "pacf"), ...) {
  params <- match.call(expand.dots = TRUE)
  params <- as.list(params)[-1]
  params[["class"]] <- NULL   # remove 'class' so it is not passed to acf()/pacf()
  if (class == "acf") {
    acf_values <- do.call(acf, c(params, list(plot = FALSE)))
  } else {
    acf_values <- do.call(pacf, c(params, list(plot = FALSE)))
  }
  acf_data <- data.frame(Lag = as.numeric(acf_values$lag),
                         ACF = as.numeric(acf_values$acf))
  seasonal_lags <- acf_data$Lag %% 1 == 0
  if (class == "acf") {
    do.call(acf, c(params, list(plot = TRUE)))
  } else {
    do.call(pacf, c(params, list(plot = TRUE)))
  }
  for (i in which(seasonal_lags)) {
    segments(x0 = acf_data$Lag[i], y0 = 0,
             x1 = acf_data$Lag[i], y1 = acf_data$ACF[i], col = "red")
  }
}
seasonal_acf  <- function(...) helper(class = "acf",  ...)
seasonal_pacf <- function(...) helper(class = "pacf", ...)


# =============================================================================
```

## 1. Data Loading and Preparation

```{r data, echo=TRUE}
# 1.  DATA LOADING & PREPARATION
# =============================================================================

raw <- read.csv("multiTimeline_AI_Growth.csv", skip = 1, header = TRUE)
colnames(raw) <- c("Month", "AI_Interest")
head(raw)
tail(raw)
dim(raw)   # 269 observations

# Convert to ts object: monthly, starting January 2004
ai.ts <- ts(raw$AI_Interest, start = c(2004, 1), frequency = 12)
ai.ts


# =============================================================================
```

## 2. Descriptive Analysis

```{r descriptive, echo=TRUE, fig.width=10, fig.height=6}
# 2.  DESCRIPTIVE ANALYSIS
# =============================================================================

# --- 2.1 Time series plot ---
plot(ai.ts, type = "o",
     ylab = "Search Interest Index (0–100)",
     xlab = "Year",
     main = "Figure 1. Monthly Global Search Interest in 'Artificial Intelligence'\n(Google Trends, Jan 2004 – May 2026)")

# Add reference line at ChatGPT launch (Nov 2022)
abline(v = 2022 + 10/12, col = "red", lty = 2)
legend("topleft", legend = c("AI Interest", "ChatGPT Launch (Nov 2022)"),
       col = c("black", "red"), lty = c(1, 2), pch = c(1, NA))

# --- 2.2 Seasonal plot (month labels) ---
par(mar = c(5, 5, 4, 2))
plot(ai.ts, type = "l",
     ylab = "Search Interest Index (0–100)",
     xlab = "Year",
     main = "Figure 2. Monthly Search Interest with Seasonal Labels")
par(mar = c(5, 4, 4, 2))
points(y = ai.ts, x = time(ai.ts), pch = as.vector(season(ai.ts)))

# --- 2.3 Summary statistics ---
summary(ai.ts)

# --- 2.4 Lag-1 scatter plot & correlation ---
y     <- ai.ts
x     <- zlag(ai.ts)
index <- 2:length(x)
cat("Lag-1 correlation:", cor(y[index], x[index]), "\n")
plot(y[index], x[index],
     xlab = "Search Interest (t-1)",
     ylab = "Search Interest (t)",
     main = "Figure 3. Lag-1 Scatter Plot")

# --- 2.5 ACF & PACF of raw series ---
par(mfrow = c(1, 2), mar = c(5, 5, 4, 2))
seasonal_acf(ai.ts,  lag.max = 48,
             main = "Figure 4. ACF of AI Search Interest",
             xlab = "Lag")
seasonal_pacf(ai.ts, lag.max = 48,
              main = "Figure 5. PACF of AI Search Interest",
              xlab = "Lag")
par(mfrow = c(1, 1), mar = c(5, 4, 4, 2))
# Slow decay in ACF → clear non-stationarity


# =============================================================================
```

## 3. Trend Models (Regression)

```{r trend, echo=TRUE, fig.width=12, fig.height=8}
# 3.  TREND MODELS (REGRESSION)
# =============================================================================

t  <- time(ai.ts)
t2 <- t^2
t3 <- t^3

# --- Model 1: Linear trend ---
model.lin <- lm(ai.ts ~ t)
summary(model.lin)
par(mar = c(5, 5, 4, 2))
plot(ai.ts, type = "o",
     ylab = "Search Interest",
     xlab = "Year",
     main = "Figure 6. Linear Trend Model")
abline(model.lin, col = "red")
par(mar = c(5, 4, 4, 2))
layout(matrix(1:4, nrow=2, ncol=2))
par(mar = c(4, 4, 3, 1))
res.tmp <- rstudent(model.lin)
plot(res.tmp, x = as.vector(time(ai.ts)), type = "l",
     xlab = "Time", ylab = "Standardised Residuals",
     main = "Residuals – Linear Model")
hist(res.tmp, xlab = "Standardised Residuals",
     main = "Histogram – Linear Model")
qqnorm(res.tmp, main = "QQ Plot – Linear Model")
qqline(res.tmp, col = 2)
acf(as.numeric(res.tmp), main = "ACF – Linear Model Residuals")
layout(1)
par(mar = c(5, 4, 4, 2) + 0.1)
shapiro.test(res.tmp)

# --- Model 2: Quadratic trend ---
model.quad <- lm(ai.ts ~ t + t2)
summary(model.quad)
par(mar = c(5, 5, 4, 2))
plot(ts(fitted(model.quad), start = c(2004, 1), frequency = 12),
     col = "red", lty = 2, type = "l",
     ylim = range(c(fitted(model.quad), as.vector(ai.ts))),
     ylab = "Search Interest",
     xlab = "Year",
     main = "Figure 7. Quadratic Trend Model")
lines(as.vector(ai.ts), type = "o")
par(mar = c(5, 4, 4, 2))
layout(matrix(1:4, nrow=2, ncol=2))
par(mar = c(4, 4, 3, 1))
res.tmp <- rstudent(model.quad)
plot(res.tmp, x = as.vector(time(ai.ts)), type = "l",
     xlab = "Time", ylab = "Standardised Residuals",
     main = "Residuals – Quadratic Model")
hist(res.tmp, xlab = "Standardised Residuals",
     main = "Histogram – Quadratic Model")
qqnorm(res.tmp, main = "QQ Plot – Quadratic Model")
qqline(res.tmp, col = 2)
acf(as.numeric(res.tmp), main = "ACF – Quadratic Model Residuals")
layout(1)
par(mar = c(5, 4, 4, 2) + 0.1)
shapiro.test(res.tmp)

# --- Model 3: Cubic trend ---
model.cub <- lm(ai.ts ~ t + t2 + t3)
summary(model.cub)
par(mar = c(5, 5, 4, 2))
plot(ts(fitted(model.cub), start = c(2004, 1), frequency = 12),
     col = "red", lty = 2, type = "l",
     ylim = range(c(fitted(model.cub), as.vector(ai.ts))),
     ylab = "Search Interest",
     xlab = "Year",
     main = "Figure 8. Cubic Trend Model")
lines(as.vector(ai.ts), type = "o")
par(mar = c(5, 4, 4, 2))
layout(matrix(1:4, nrow=2, ncol=2))
par(mar = c(4, 4, 3, 1))
res.tmp <- rstudent(model.cub)
plot(res.tmp, x = as.vector(time(ai.ts)), type = "l",
     xlab = "Time", ylab = "Standardised Residuals",
     main = "Residuals – Cubic Model")
hist(res.tmp, xlab = "Standardised Residuals",
     main = "Histogram – Cubic Model")
qqnorm(res.tmp, main = "QQ Plot – Cubic Model")
qqline(res.tmp, col = 2)
acf(as.numeric(res.tmp), main = "ACF – Cubic Model Residuals")
layout(1)
par(mar = c(5, 4, 4, 2) + 0.1)
shapiro.test(res.tmp)

# --- Model 4: Seasonal means model ---
month. <- season(ai.ts)
model.seas <- lm(ai.ts ~ month. - 1)
summary(model.seas)
layout(matrix(1:4, nrow=2, ncol=2))
par(mar = c(4, 4, 3, 1))
res.tmp <- rstudent(model.seas)
plot(res.tmp, x = as.vector(time(ai.ts)), type = "l",
     xlab = "Time", ylab = "Standardised Residuals",
     main = "Residuals – Seasonal Model")
hist(res.tmp, xlab = "Standardised Residuals",
     main = "Histogram – Seasonal Model")
qqnorm(res.tmp, main = "QQ Plot – Seasonal Model")
qqline(res.tmp, col = 2)
acf(as.numeric(res.tmp), main = "ACF – Seasonal Model Residuals")
layout(1)
par(mar = c(5, 4, 4, 2) + 0.1)
shapiro.test(res.tmp)

# --- Model 5: Seasonal + quadratic ---
model.seas.quad <- lm(ai.ts ~ month. + t + t2 - 1)
summary(model.seas.quad)
par(mar = c(5, 5, 4, 2))
plot(ts(fitted(model.seas.quad), start = c(2004, 1), frequency = 12),
     col = "red", lty = 2, type = "l",
     ylim = range(c(fitted(model.seas.quad), as.vector(ai.ts))),
     ylab = "Search Interest",
     xlab = "Year",
     main = "Figure 9. Seasonal + Quadratic Trend Model")
lines(as.vector(ai.ts), type = "o")
par(mar = c(5, 4, 4, 2))
layout(matrix(1:4, nrow=2, ncol=2))
par(mar = c(4, 4, 3, 1))
res.tmp <- rstudent(model.seas.quad)
plot(res.tmp, x = as.vector(time(ai.ts)), type = "l",
     xlab = "Time", ylab = "Standardised Residuals",
     main = "Residuals – Seasonal + Quadratic Model")
hist(res.tmp, xlab = "Standardised Residuals",
     main = "Histogram – Seasonal + Quadratic Model")
qqnorm(res.tmp, main = "QQ Plot – Seasonal + Quadratic Model")
qqline(res.tmp, col = 2)
acf(as.numeric(res.tmp), main = "ACF – Seasonal + Quadratic Model Residuals")
layout(1)
par(mar = c(5, 4, 4, 2) + 0.1)
shapiro.test(res.tmp)

# --- Model 6: Harmonic ---
har.       <- harmonic(ai.ts, 1)
data.har   <- data.frame(ai.ts, har.)
model.har  <- lm(ai.ts ~ cos.2.pi.t. + sin.2.pi.t., data = data.har)
summary(model.har)
layout(matrix(1:4, nrow=2, ncol=2))
par(mar = c(4, 4, 3, 1))
res.tmp <- rstudent(model.har)
plot(res.tmp, x = as.vector(time(ai.ts)), type = "l",
     xlab = "Time", ylab = "Standardised Residuals",
     main = "Residuals – Harmonic Model")
hist(res.tmp, xlab = "Standardised Residuals",
     main = "Histogram – Harmonic Model")
qqnorm(res.tmp, main = "QQ Plot – Harmonic Model")
qqline(res.tmp, col = 2)
acf(as.numeric(res.tmp), main = "ACF – Harmonic Model Residuals")
layout(1)
par(mar = c(5, 4, 4, 2) + 0.1)
shapiro.test(res.tmp)

# --- AIC / BIC comparison for trend models ---
sort.score(AIC(model.lin, model.quad, model.cub,
               model.seas, model.seas.quad, model.har), score = "aic")
sort.score(AIC(model.lin, model.quad, model.cub,
               model.seas, model.seas.quad, model.har,
               k = log(length(ai.ts))), score = "aic")

# --- Forecasts from best trend model (seasonal + quadratic) ---
h              <- 10   # 10 months ahead
last.t         <- t[length(t)]
dataFreq       <- frequency(ai.ts)
new.t          <- seq(last.t + 1/dataFreq,
                      last.t + h/dataFreq, by = 1/dataFreq)
ahead.seas.quad <- data.frame(
  month. = c("June", "July", "August", "September", "October",
             "November", "December", "January", "February", "March"),
  t  = new.t,
  t2 = new.t^2
)
frc.seas.quad <- predict(model.seas.quad,
                         newdata = ahead.seas.quad,
                         interval = "prediction")

par(mar = c(5, 5, 4, 2))
plot(ai.ts,
     xlim = c(2004, last.t + h/dataFreq + 0.1),
     ylim = c(0, 130),
     ylab = "Search Interest Index",
     xlab = "Year",
     main = "Figure 10. 10-Month Forecasts – Seasonal + Quadratic Trend Model")
lines(ts(fitted(model.seas.quad), start = c(2004,1), frequency = 12),
      col = "green")
lines(ts(as.vector(frc.seas.quad[, 1]), start = c(2026, 6), frequency = 12),
      col = "red")
lines(ts(as.vector(frc.seas.quad[, 2]), start = c(2026, 6), frequency = 12),
      col = "blue")
lines(ts(as.vector(frc.seas.quad[, 3]), start = c(2026, 6), frequency = 12),
      col = "blue")
legend("topleft", lty = 1,
       col = c("black", "green", "red", "blue"),
       legend = c("Data", "Fitted", "Forecasts", "95% PI"))
par(mar = c(5, 4, 4, 2))


# =============================================================================
```

## 4. ARIMA Modelling

```{r arima, echo=TRUE, fig.width=12, fig.height=10}
# 4.  ARIMA MODELLING
# =============================================================================

# --- 4.1 Normality & Box-Cox transformation ---
par(mar = c(5, 5, 4, 2))
qqnorm(ai.ts,
       main = "Figure 11. QQ Plot – Raw Series",
       xlab = "Theoretical Quantiles",
       ylab = "Sample Quantiles")
qqline(ai.ts, col = 2)
par(mar = c(5, 4, 4, 2))
shapiro.test(ai.ts)

# BoxCox.ar from TSA — professor's method
BC    <- BoxCox.ar(ai.ts)
BC$ci
lambda <- BC$lambda[which(max(BC$loglike) == BC$loglike)]
lambda
BC.ai <- (ai.ts^lambda - 1) / lambda

par(mar = c(5, 5, 4, 2))
plot(BC.ai, type = "o",
     ylim = c(floor(min(BC.ai) * 10) / 10 - 0.05,
              ceiling(max(BC.ai) * 10) / 10 + 0.05),
     ylab = "BC Transformed Index",
     xlab = "Year",
     main = "Figure 12. Box-Cox Transformed Series")
par(mar = c(5, 4, 4, 2))
par(mar = c(5, 5, 4, 2))
par(mar = c(5, 5, 4, 2))
qqnorm(BC.ai,
       main = "Figure 13. QQ Plot – BC Transformed Series",
       xlab = "Theoretical Quantiles",
       ylab = "Sample Quantiles")
qqline(BC.ai, col = 2)
par(mar = c(5, 4, 4, 2))
par(mar = c(5, 4, 4, 2))
shapiro.test(BC.ai)

# --- 4.2 Stationarity tests on raw series ---
adf.test(ai.ts)
pp.test(ai.ts)

# --- 4.3 First difference ---
diff1.ai <- diff(ai.ts, differences = 1)
par(mar = c(5, 5, 4, 2))
plot(diff1.ai, type = "o",
     ylim = c(min(diff1.ai) - 2, max(diff1.ai) + 2),
     ylab = "First Difference",
     xlab = "Year",
     main = "Figure 14. First Difference of AI Search Interest")
par(mar = c(5, 4, 4, 2))

adf.test(diff1.ai)
pp.test(diff1.ai)
kpss.test(diff1.ai)

par(mfrow = c(1, 2), mar = c(5, 5, 4, 2))
acf(diff1.ai,  lag.max = 48,
    main = "Figure 15. ACF – First Difference",
    xlab = "Lag")
pacf(diff1.ai, lag.max = 48,
     main = "Figure 16. PACF – First Difference",
     xlab = "Lag")
par(mfrow = c(1, 1), mar = c(5, 4, 4, 2))

# --- 4.4 Second difference (if needed) ---
diff2.ai <- diff(ai.ts, differences = 2)
par(mar = c(5, 5, 4, 2))
plot(diff2.ai, type = "o",
     ylim = c(min(diff2.ai) - 2, max(diff2.ai) + 2),
     ylab = "Second Difference",
     xlab = "Year",
     main = "Figure 17. Second Difference of AI Search Interest")
par(mar = c(5, 4, 4, 2))

adf.test(diff2.ai)
pp.test(diff2.ai)
kpss.test(diff2.ai)

par(mfrow = c(1, 2), mar = c(5, 5, 4, 2))
acf(diff2.ai,  lag.max = 48,
    main = "Figure 18. ACF – Second Difference",
    xlab = "Lag")
pacf(diff2.ai, lag.max = 48,
     main = "Figure 19. PACF – Second Difference",
     xlab = "Lag")
par(mfrow = c(1, 1), mar = c(5, 4, 4, 2))

# d = 1 (use diff1.ai for model identification)

# --- 4.5 EACF ---
eacf(diff1.ai, ar.max = 6, ma.max = 6)

# --- 4.6 BIC subset selection ---
res.arma <- armasubsets(y = diff1.ai, nar = 6, nma = 6,
                         y.name = "p", ar.method = "ols")
plot(res.arma)

# Candidate models (update based on your ACF/PACF/EACF output):
# {ARIMA(0,1,1), ARIMA(1,1,1), ARIMA(2,1,1),
#  ARIMA(0,1,2), ARIMA(1,1,2), ARIMA(2,1,2)}

# --- 4.7 Fit candidate ARIMA models ---

# ARIMA(0,1,1)
m.011.css <- Arima(ai.ts, order = c(0,1,1), method = "CSS")
m.011.ml  <- tryCatch(
  Arima(ai.ts, order = c(0,1,1), method = "ML"),
  error = function(e) tryCatch(
    Arima(ai.ts, order = c(0,1,1), method = "CSS-ML"),
    error = function(e) Arima(ai.ts, order = c(0,1,1), method = "CSS")
  )
)
coeftest(m.011.ml)
residual.analysis(m.011.ml, lag.max = 30)

# ARIMA(1,1,1)
m.111.css <- Arima(ai.ts, order = c(1,1,1), method = "CSS")
m.111.ml  <- tryCatch(
  Arima(ai.ts, order = c(1,1,1), method = "ML"),
  error = function(e) tryCatch(
    Arima(ai.ts, order = c(1,1,1), method = "CSS-ML"),
    error = function(e) Arima(ai.ts, order = c(1,1,1), method = "CSS")
  )
)
coeftest(m.111.ml)
residual.analysis(m.111.ml, lag.max = 30)

# ARIMA(2,1,1)
m.211.css <- Arima(ai.ts, order = c(2,1,1), method = "CSS")
m.211.ml  <- tryCatch(
  Arima(ai.ts, order = c(2,1,1), method = "ML"),
  error = function(e) tryCatch(
    Arima(ai.ts, order = c(2,1,1), method = "CSS-ML"),
    error = function(e) Arima(ai.ts, order = c(2,1,1), method = "CSS")
  )
)
coeftest(m.211.ml)
residual.analysis(m.211.ml, lag.max = 30)

# ARIMA(0,1,2)
m.012.css <- Arima(ai.ts, order = c(0,1,2), method = "CSS")
m.012.ml  <- tryCatch(
  Arima(ai.ts, order = c(0,1,2), method = "ML"),
  error = function(e) tryCatch(
    Arima(ai.ts, order = c(0,1,2), method = "CSS-ML"),
    error = function(e) Arima(ai.ts, order = c(0,1,2), method = "CSS")
  )
)
coeftest(m.012.ml)
residual.analysis(m.012.ml, lag.max = 30)

# ARIMA(1,1,2)
m.112.css <- Arima(ai.ts, order = c(1,1,2), method = "CSS")
m.112.ml  <- tryCatch(
  Arima(ai.ts, order = c(1,1,2), method = "ML"),
  error = function(e) tryCatch(
    Arima(ai.ts, order = c(1,1,2), method = "CSS-ML"),
    error = function(e) Arima(ai.ts, order = c(1,1,2), method = "CSS")
  )
)
coeftest(m.112.ml)
residual.analysis(m.112.ml, lag.max = 30)

# ARIMA(2,1,2)
m.212.css <- Arima(ai.ts, order = c(2,1,2), method = "CSS")
m.212.ml  <- tryCatch(
  Arima(ai.ts, order = c(2,1,2), method = "ML"),
  error = function(e) tryCatch(
    Arima(ai.ts, order = c(2,1,2), method = "CSS-ML"),
    error = function(e) Arima(ai.ts, order = c(2,1,2), method = "CSS")
  )
)
coeftest(m.212.ml)
residual.analysis(m.212.ml, lag.max = 30)

# --- 4.8 AIC / BIC comparison ---
safe.AIC(m.011.ml, m.111.ml, m.211.ml, m.012.ml, m.112.ml, m.212.ml)
safe.BIC(m.011.ml, m.111.ml, m.211.ml, m.012.ml, m.112.ml, m.212.ml)

# --- 4.9 Accuracy table (CSS fits) ---
Sm.011 <- accuracy(m.011.css)[1:7]
Sm.111 <- accuracy(m.111.css)[1:7]
Sm.211 <- accuracy(m.211.css)[1:7]
Sm.012 <- accuracy(m.012.css)[1:7]
Sm.112 <- accuracy(m.112.css)[1:7]
Sm.212 <- accuracy(m.212.css)[1:7]
df.arima <- data.frame(rbind(Sm.011, Sm.111, Sm.211,
                              Sm.012, Sm.112, Sm.212))
colnames(df.arima)  <- c("ME","RMSE","MAE","MPE","MAPE","MASE","ACF1")
rownames(df.arima)  <- c("ARIMA(0,1,1)","ARIMA(1,1,1)","ARIMA(2,1,1)",
                          "ARIMA(0,1,2)","ARIMA(1,1,2)","ARIMA(2,1,2)")
round(df.arima, digits = 3)

# --- 4.10 Overparameterisation check (replace with your best model) ---
# Example: if ARIMA(0,1,1) is best, check ARIMA(1,1,1) and ARIMA(0,1,2)
coeftest(m.111.ml)   # AR(1) significant?
coeftest(m.012.ml)   # MA(2) significant?

# --- 4.11 ARIMA Forecasts (10 months ahead) ---
# Replace m.011.ml with your selected best ARIMA model
frc.arima <- forecast(m.011.ml, h = 10)
par(mar = c(5, 5, 5, 2))
plot(frc.arima,
     main = "Figure 20. 10-Month Forecasts – ARIMA Model",
     ylab = "Search Interest Index",
     xlab = "Year")
par(mar = c(5, 4, 4, 2))
frc.arima


# =============================================================================
```

## 5. SARIMA Modelling

```{r sarima, echo=TRUE, fig.width=12, fig.height=10}
# 5.  SARIMA MODELLING
# =============================================================================

# --- 5.1 Seasonal ACF/PACF of raw series ---
par(mfrow = c(1, 2), mar = c(5, 5, 4, 2))
seasonal_acf(ai.ts,  lag.max = 60,
             main = "Figure 21. Seasonal ACF – AI Search Interest",
             xlab = "Lag")
seasonal_pacf(ai.ts, lag.max = 60,
              main = "Figure 22. Seasonal PACF – AI Search Interest",
              xlab = "Lag")
par(mfrow = c(1, 1), mar = c(5, 4, 4, 2))

# --- 5.2 Step 1: Apply seasonal difference D=1, s=12 ---
m.s1 <- Arima(ai.ts, order = c(0,0,0),
              seasonal = list(order = c(0,1,0), period = 12))
res.s1 <- residuals(m.s1)
par(mar = c(5, 5, 4, 2))
plot(res.s1, type = "o",
     ylab = "Residuals",
     xlab = "Year",
     main = "Figure 23. Residuals after Seasonal Difference")
par(mfrow = c(1, 2), mar = c(5, 5, 4, 2))
seasonal_acf(res.s1,  lag.max = 48, main = "ACF after D=1", xlab = "Lag")
seasonal_pacf(res.s1, lag.max = 48, main = "PACF after D=1", xlab = "Lag")
par(mfrow = c(1, 1), mar = c(5, 4, 4, 2))

# --- 5.3 Step 2: Add seasonal MA(1) ---
m.s2 <- Arima(ai.ts, order = c(0,0,0),
              seasonal = list(order = c(0,1,1), period = 12))
res.s2 <- residuals(m.s2)
par(mar = c(5, 5, 4, 2))
plot(res.s2, type = "o",
     ylab = "Residuals",
     xlab = "Year",
     main = "Figure 24. Residuals after SARMA(0,1,1)_12")
par(mfrow = c(1, 2), mar = c(5, 5, 4, 2))
seasonal_acf(res.s2,  lag.max = 48, main = "ACF – SARMA(0,1,1)_12 residuals", xlab = "Lag")
seasonal_pacf(res.s2, lag.max = 48, main = "PACF – SARMA(0,1,1)_12 residuals", xlab = "Lag")
par(mfrow = c(1, 1), mar = c(5, 4, 4, 2))

# --- 5.4 Step 3: Add ordinary difference d=1 ---
m.s3 <- Arima(ai.ts, order = c(0,1,0),
              seasonal = list(order = c(0,1,1), period = 12))
res.s3 <- residuals(m.s3)
par(mar = c(5, 5, 4, 2))
plot(res.s3, type = "o",
     ylab = "Residuals",
     xlab = "Year",
     main = "Figure 25. Residuals after d=1, D=1")
par(mfrow = c(1, 2), mar = c(5, 5, 4, 2))
seasonal_acf(res.s3,  lag.max = 48, main = "ACF – d=1, D=1 residuals", xlab = "Lag")
seasonal_pacf(res.s3, lag.max = 48, main = "PACF – d=1, D=1 residuals", xlab = "Lag")
par(mfrow = c(1, 1), mar = c(5, 4, 4, 2))
adf.test(res.s3)
pp.test(res.s3)

# --- 5.5 Identify ordinary ARMA orders from residuals ---
eacf(res.s3, ar.max = 5, ma.max = 5)
layout(1)
bic.sarima <- armasubsets(y = res.s3, nar = 5, nma = 5,
                           y.name = "p", ar.method = "ols")
plot(bic.sarima)

# Candidate SARIMA models:
# {SARIMA(0,1,1)x(0,1,1)_12,
#  SARIMA(1,1,1)x(0,1,1)_12,
#  SARIMA(0,1,2)x(0,1,1)_12,
#  SARIMA(1,1,2)x(0,1,1)_12}

# --- 5.6 Fit candidate SARIMA models ---

# SARIMA(0,1,1)x(0,1,1)_12
sm.011_011.ml  <- tryCatch(
  Arima(ai.ts, order = c(0,1,1),
        seasonal = list(order = c(0,1,1), period = 12), method = "ML"),
  error = function(e) tryCatch(
    Arima(ai.ts, order = c(0,1,1),
          seasonal = list(order = c(0,1,1), period = 12), method = "CSS-ML"),
    error = function(e) Arima(ai.ts, order = c(0,1,1),
          seasonal = list(order = c(0,1,1), period = 12), method = "CSS")
  )
)
sm.011_011.css <- Arima(ai.ts, order = c(0,1,1),
                         seasonal = list(order = c(0,1,1), period = 12),
                         method = "CSS")
coeftest(sm.011_011.ml)
residual.analysis(sm.011_011.ml, lag.max = 30, class = "SARIMA")

# SARIMA(1,1,1)x(0,1,1)_12
sm.111_011.ml  <- tryCatch(
  Arima(ai.ts, order = c(1,1,1),
        seasonal = list(order = c(0,1,1), period = 12), method = "ML"),
  error = function(e) tryCatch(
    Arima(ai.ts, order = c(1,1,1),
          seasonal = list(order = c(0,1,1), period = 12), method = "CSS-ML"),
    error = function(e) Arima(ai.ts, order = c(1,1,1),
          seasonal = list(order = c(0,1,1), period = 12), method = "CSS")
  )
)
sm.111_011.css <- Arima(ai.ts, order = c(1,1,1),
                         seasonal = list(order = c(0,1,1), period = 12),
                         method = "CSS")
coeftest(sm.111_011.ml)
residual.analysis(sm.111_011.ml, lag.max = 30, class = "SARIMA")

# SARIMA(0,1,2)x(0,1,1)_12
sm.012_011.ml  <- tryCatch(
  Arima(ai.ts, order = c(0,1,2),
        seasonal = list(order = c(0,1,1), period = 12), method = "ML"),
  error = function(e) tryCatch(
    Arima(ai.ts, order = c(0,1,2),
          seasonal = list(order = c(0,1,1), period = 12), method = "CSS-ML"),
    error = function(e) Arima(ai.ts, order = c(0,1,2),
          seasonal = list(order = c(0,1,1), period = 12), method = "CSS")
  )
)
sm.012_011.css <- Arima(ai.ts, order = c(0,1,2),
                         seasonal = list(order = c(0,1,1), period = 12),
                         method = "CSS")
coeftest(sm.012_011.ml)
residual.analysis(sm.012_011.ml, lag.max = 30, class = "SARIMA")

# SARIMA(1,1,2)x(0,1,1)_12
sm.112_011.ml  <- tryCatch(
  Arima(ai.ts, order = c(1,1,2),
        seasonal = list(order = c(0,1,1), period = 12), method = "ML"),
  error = function(e) tryCatch(
    Arima(ai.ts, order = c(1,1,2),
          seasonal = list(order = c(0,1,1), period = 12), method = "CSS-ML"),
    error = function(e) Arima(ai.ts, order = c(1,1,2),
          seasonal = list(order = c(0,1,1), period = 12), method = "CSS")
  )
)
sm.112_011.css <- Arima(ai.ts, order = c(1,1,2),
                         seasonal = list(order = c(0,1,1), period = 12),
                         method = "CSS")
coeftest(sm.112_011.ml)
residual.analysis(sm.112_011.ml, lag.max = 30, class = "SARIMA")

# --- 5.7 AIC / BIC comparison ---
safe.AIC(sm.011_011.ml, sm.111_011.ml, sm.012_011.ml, sm.112_011.ml)
safe.BIC(sm.011_011.ml, sm.111_011.ml, sm.012_011.ml, sm.112_011.ml)

# --- 5.8 Accuracy table ---
Ssm.011_011 <- accuracy(sm.011_011.css)[1:7]
Ssm.111_011 <- accuracy(sm.111_011.css)[1:7]
Ssm.012_011 <- accuracy(sm.012_011.css)[1:7]
Ssm.112_011 <- accuracy(sm.112_011.css)[1:7]
df.sarima <- data.frame(rbind(Ssm.011_011, Ssm.111_011,
                               Ssm.012_011, Ssm.112_011))
colnames(df.sarima) <- c("ME","RMSE","MAE","MPE","MAPE","MASE","ACF1")
rownames(df.sarima) <- c("SARIMA(0,1,1)x(0,1,1)_12",
                          "SARIMA(1,1,1)x(0,1,1)_12",
                          "SARIMA(0,1,2)x(0,1,1)_12",
                          "SARIMA(1,1,2)x(0,1,1)_12")
round(df.sarima, digits = 3)

# --- 5.9 Overparameterisation check ---
# Example: if SARIMA(0,1,1)x(0,1,1)_12 is best:
sm.011_011.over1 <- tryCatch(
  Arima(ai.ts, order = c(1,1,1),
        seasonal = list(order = c(0,1,1), period = 12), method = "ML"),
  error = function(e) tryCatch(
    Arima(ai.ts, order = c(1,1,1),
          seasonal = list(order = c(0,1,1), period = 12), method = "CSS-ML"),
    error = function(e) Arima(ai.ts, order = c(1,1,1),
          seasonal = list(order = c(0,1,1), period = 12), method = "CSS")
  )
)
coeftest(sm.011_011.over1)   # AR(1) significant?

sm.011_011.over2 <- tryCatch(
  Arima(ai.ts, order = c(0,1,2),
        seasonal = list(order = c(0,1,1), period = 12), method = "ML"),
  error = function(e) tryCatch(
    Arima(ai.ts, order = c(0,1,2),
          seasonal = list(order = c(0,1,1), period = 12), method = "CSS-ML"),
    error = function(e) Arima(ai.ts, order = c(0,1,2),
          seasonal = list(order = c(0,1,1), period = 12), method = "CSS")
  )
)
coeftest(sm.011_011.over2)   # MA(2) significant?

# --- 5.10 SARIMA Forecasts ---
# Replace sm.011_011.ml with your selected best SARIMA model
frc.sarima <- forecast(sm.011_011.ml, h = 10)
par(mar = c(5, 5, 5, 2))
plot(frc.sarima,
     main = "Figure 26. 10-Month Forecasts – SARIMA Model",
     ylab = "Search Interest Index",
     xlab = "Year")
par(mar = c(5, 4, 4, 2))
frc.sarima


# =============================================================================
```

## 6. GARCH / ARMA+GARCH Modelling

```{r garch, echo=TRUE, fig.width=12, fig.height=10}
# 6.  GARCH / ARMA+GARCH MODELLING
# =============================================================================

# --- 6.1 Compute returns (first difference of log — use for GARCH) ---
# Because ai.ts has zeros, shift by +1 before log
ai.shifted  <- ai.ts + 1
ai.returns  <- diff(log(ai.shifted))

par(mar = c(5, 5, 4, 2))
plot(ai.returns, type = "o",
     ylab = "Log Returns",
     xlab = "Year",
     main = "Figure 27. Returns of AI Search Interest (diff of log)")
par(mar = c(5, 4, 4, 2))

# --- 6.2 McLeod-Li test for ARCH effects ---
par(mar = c(5, 5, 4, 2))
McLeod.Li.test(y = ai.returns,
               main = "Figure 28. McLeod-Li Test – AI Returns")
par(mar = c(5, 4, 4, 2))
# Significant lags → ARCH effects present → GARCH needed

# --- 6.3 ACF/PACF of returns ---
par(mfrow = c(1, 2), mar = c(5, 5, 4, 2))
acf(ai.returns,  main = "Figure 29. ACF of Returns", xlab = "Lag")
pacf(ai.returns, main = "Figure 30. PACF of Returns", xlab = "Lag")
par(mfrow = c(1, 1), mar = c(5, 4, 4, 2))

adf.test(ai.returns)   # Should be stationary

# --- 6.4 ARMA specification for mean equation ---
eacf(ai.returns, ar.max = 5, ma.max = 5)
res.arma2 <- armasubsets(y = ai.returns, nar = 5, nma = 5,
                          y.name = "p", ar.method = "ols")
plot(res.arma2)

# --- 6.5 Fit ARMA models for mean equation ---
# (Keep this simple — the GARCH part handles variance)
ma.011.css <- Arima(ai.returns, order = c(0,0,1), method = "CSS")
ma.011.ml  <- tryCatch(
  Arima(ai.returns, order = c(0,0,1), method = "ML"),
  error = function(e) tryCatch(
    Arima(ai.returns, order = c(0,0,1), method = "CSS-ML"),
    error = function(e) Arima(ai.returns, order = c(0,0,1), method = "CSS")
  )
)
coeftest(ma.011.ml)
residual.analysis(ma.011.ml, lag.max = 30)

ma.111.css <- Arima(ai.returns, order = c(1,0,1), method = "CSS")
ma.111.ml  <- tryCatch(
  Arima(ai.returns, order = c(1,0,1), method = "ML"),
  error = function(e) tryCatch(
    Arima(ai.returns, order = c(1,0,1), method = "CSS-ML"),
    error = function(e) Arima(ai.returns, order = c(1,0,1), method = "CSS")
  )
)
coeftest(ma.111.ml)
residual.analysis(ma.111.ml, lag.max = 30)

safe.AIC(ma.011.ml, ma.111.ml)
safe.BIC(ma.011.ml, ma.111.ml)

# --- 6.6 Check ARCH effects in best ARMA residuals ---
# Replace ma.011.ml with your best ARMA model
best.arma.res    <- ma.011.ml$residuals
abs.res          <- abs(best.arma.res)
sq.res           <- best.arma.res^2

par(mfrow = c(1, 2), mar = c(5, 5, 4, 2))
acf(abs.res,  ci.type = "ma", main = "Figure 31. ACF – |Residuals|", xlab = "Lag")
pacf(abs.res, main = "Figure 32. PACF – |Residuals|", xlab = "Lag")
par(mfrow = c(1, 1), mar = c(5, 4, 4, 2))
eacf(abs.res, ar.max = 5, ma.max = 5)

par(mfrow = c(1, 2), mar = c(5, 5, 4, 2))
acf(sq.res,  ci.type = "ma", main = "Figure 33. ACF – Residuals²", xlab = "Lag")
pacf(sq.res, main = "Figure 34. PACF – Residuals²", xlab = "Lag")
par(mfrow = c(1, 1), mar = c(5, 4, 4, 2))
eacf(sq.res, ar.max = 5, ma.max = 5)

# --- 6.7 Fit GARCH models ---
# Candidates (update from EACF): {GARCH(0,1), GARCH(1,1), GARCH(1,2)}

g.01 <- garch(best.arma.res, order = c(0,1), trace = FALSE)
summary(g.01)
residual.analysis(g.01, class = "garch")

g.11 <- garch(best.arma.res, order = c(1,1), trace = FALSE)
summary(g.11)
residual.analysis(g.11, class = "garch")

g.12 <- garch(best.arma.res, order = c(1,2), trace = FALSE)
summary(g.12)
residual.analysis(g.12, start = 3, class = "garch")

sc.AIC.garch <- AIC(g.01, g.11, g.12)
sc.BIC.garch <- AIC(g.01, g.11, g.12, k = log(length(best.arma.res)))
sort.score(sc.AIC.garch, score = "aic")
sort.score(sc.BIC.garch, score = "aic")

# --- 6.8 Fit combined ARMA+GARCH using rugarch ---
# NOTE: garchOrder = c(q, p) in rugarch (reversed from garch())
# Update armaOrder and garchOrder to match your best models above

spec.ag <- ugarchspec(
  variance.model = list(model = "sGARCH",
                        garchOrder = c(1, 1)),   # GARCH(0,1) → c(1,0); GARCH(1,1) → c(1,1)
  mean.model     = list(armaOrder = c(0, 1),     # ARMA(0,1) = MA(1)
                        include.mean = FALSE),
  distribution.model = "norm"
)
fit.ag <- ugarchfit(spec = spec.ag, data = ai.returns)
fit.ag
residual.analysis(fit.ag, class = "ARMA-GARCH")

# --- 6.9 AIC/BIC for ARMA+GARCH ---
AICs.ag <- infocriteria(fit.ag)[1]
BICs.ag <- infocriteria(fit.ag)[2]
cat("ARMA+GARCH AIC:", AICs.ag, "  BIC:", BICs.ag, "\n")

# Plot conditional variance
plot(fit.ag, which = 2)

# --- 6.10 GARCH Forecasts (10 steps ahead) ---
forc.ag <- ugarchforecast(fit.ag, data = ai.returns, n.ahead = 10)
plot(forc.ag, which = 1)   # Series forecasts
plot(forc.ag, which = 3)   # Conditional variance forecasts
forc.ag


# =============================================================================
```

## 7. Final Model Comparison

```{r comparison, echo=TRUE}
# 7.  MODEL COMPARISON SUMMARY
# =============================================================================

# Compare best model from each family
# Replace model names with your actual best models after running the above

cat("\n===== FINAL MODEL COMPARISON =====\n")

# Trend model accuracy (CSS-equivalent via lm)
cat("Best Trend Model: Seasonal + Quadratic\n")
cat("AIC:", AIC(model.seas.quad), "\n")

# ARIMA
cat("\nBest ARIMA model:\n")
# e.g. sort.score output will tell you — replace m.011.ml below
cat("AIC:", AIC(m.011.ml), "  BIC:", BIC(m.011.ml), "\n")
accuracy(m.011.css)

# SARIMA
cat("\nBest SARIMA model:\n")
cat("AIC:", AIC(sm.011_011.ml), "  BIC:", BIC(sm.011_011.ml), "\n")
accuracy(sm.011_011.css)

# ARMA+GARCH
cat("\nARMA+GARCH:\n")
cat("AIC:", infocriteria(fit.ag)[1],
    "  BIC:", infocriteria(fit.ag)[2], "\n")


```
